Over Twenty Years of Virtual Machine Development and Debugging Through Simulation

### 2018-08-15 paper.tex
added fig

\begin{figure}[bth!]
		\centering
		\includegraphics[width=0.6\linewidth]{figures/VMCompilation}
		\caption{Cog VM compilation}
		\label{fig:VMCompilation}
\end{figure}

@@

The heap is simulated as a large contiguous byte array. References between objects are effectively indexes inside the byte array instead of pointers. All the C variables are simulated as Smalltalk objects. They used specific classes, such as \emph{CArrayAccessor} over normal Smalltalk classes, to emulate the C behavior (only array accesses are available in C, not high level iterator APIs, etc.). An interesting example is the execution stack. The stack is represented in \openSmalltalkVM as a double linked list of stack pages which are maintained by the VM. Each stack page is represented as a Smalltalk object.

MEMORY FIGURE?

@@

@@

The debugging process used here is exampled in a youtube video \footnote{https://www.youtube.com/watch?v=hctMBGAXVSs}.

>>>>

\begin{figure}[bth!]
		\centering
		\includegraphics[width=0.55\linewidth]{figures/VMCompilation}
		\caption{Cog VM compilation}
		\label{fig:VMCompilation}
\end{figure}

@@

Figure \ref{fig:SimulatedMemory} describes, from a high-level perspective, the memory used by \openSmalltalkVM at runtime and the simulated counter-part. Let's detail briefly the memory used by the production VM, on the top of the figure. On the left side, which is usually the low addresses, we can find Text, the memory section holding the native code of the VM (the interpreter, the memory manager and the JIT itself, but not the code compiled by the JIT). Then, Data holds initialized and uninitialized data, mainly the C variables used globally in the VM. On higher addresses, we can find the beginning of the memory managed by the VM. At start-up, the VM mmaps a memory region used for the executable code generated by the JIT (optional executable section, exists only if the JIT is enabled), for the young objects and for the old objects present in the snapshot started plus a little more space for the first tenures\footnote{A tenure is the process of promoting young objects to old objects in a generational GC.} to be performed without triggering the full garbage collector. Later during execution, as old space grows, new memory regions are mmap'ed on higher addresses to store other old objects. Usually at very high addresses, we can find the C stack. Room in the C stack is alloca'ed at VM start-up to hold the Smalltalk stack. Both stacks are disjointed and managed differently.

In the simulator, the heap is simulated as a large contiguous byte array. References between objects are effectively indexes inside the byte array instead of pointers. All the Slang variables, normally translated to C variabled, are simulated as Smalltalk objects. They used specific classes, such as \emph{CArrayAccessor} over normal Smalltalk classes, to emulate the C behavior (only array accesses are available in C, not high level iterator APIs, etc.). The Slang code is executed as Smalltalk code. The Smalltalk stack is represented in \openSmalltalkVM as a double linked list of stack pages which are maintained by the VM. Each stack page is represented as a Smalltalk object in the simulator.

\begin{figure*}[bth!]
		\centering
		\includegraphics[width=0.9\linewidth]{figures/SimulatedMemory}
		\caption{Runtime Memory and Simulated Counter-Part}
		\label{fig:SimulatedMemory}
\end{figure*}

@@

IN-IMAGE COMP FIG?

@@

The debugging process used here is exampled in a youtube video\footnote{https://www.youtube.com/watch?v=hctMBGAXVSs}.
